---
title: "Data Science in Psychology"
author: "Team Psychology"
toc: true
number-sections: true
highlight-style: pygments
format: 
  html:  
    code-fold: true
    html-math-method: katex
  pdf: 
    geometry: 
      - top=30mm
      - left=20mm
  docx: default
---
# Authors

- Sashank Gangadharabhotla
- Brahmendra Charan Attanti
- Yamini Priyanka Kandikattu 
- Anila Reddy Musku
- Megha Veeregowda


# What is Psychology 
The scientific study of behavior and the mind is known as psychology. In simple words, it explains how the human brain works and why it behaves in such a particular way.

# Question of interest?
Our main aim is to find out the factors leading to psychological issues like depression (mainly focusing on the Age and Gender of that person) and it is important to figure out those factors as depression strongly effects human's life.

Source - we got the data from kaggle 
https://www.kaggle.com/datasets/diegobabativa/depression

# Description of the dataset 

1 - Our dataset contains 1429 rows and 23 columns which is a reasonable size to perform analysis.

2 - target label is the 'depression' column which takes on 0 and 1 values with 0 being not depressed and 1 being depressed.

3 - The independent columns in our data set which are going to affect our target label are majorly of two categories - Family related information and financial perspective.

4 - some of the columns related to one's personal(family) information are Age, Gender, Married, Number of children, total numbers in the family, educational level etc..

5 - some of the columns related to one's financial information are the incoming salary, living expenses, Gained assets, buisinesses if any and so on.

# Analysing the dataset

## Loading Library
```{r}
library('tidyverse')
library('ggplot2')
```

## Reading Dataset
```{r}

url <- 'https://raw.githubusercontent.com/Charan-AB/Data-science-in-psychology/main/b_depressed.csv'
df <- readr::read_csv(url)
head(df)
```

```{r}
dim(df)
names(df)
```

```{r}
df_stat <- df
colSums(is.na(df_stat))
```

We are having just 20 null values in 'no_lasting_investmen' column - dropping those
```{r}
df <- na.omit(df)
colSums(is.na(df))
```

looking at our target label - 'depressed' column

```{r}
unique(df)

```
 0 refers to not depressed and 1 refers to being depressed
 
## Visualizations

### understanding of the data attributes by looking at how they are distributed
```{r}
plot(density(df$Age))
# most of the age between 20 and 40 mark 
```
```{r}
qqnorm(df$Age, col = "skyblue3", 
       main = "Age Q-Q Plot")
qqline(df$Age, col = "red", lwd = 3)

# by looking at the graph above and the qqplot we can say that it's slightly a bit right skewed
```
```{r}
hist(df$education_level)

```
Most of the people have studied till their 10th standard(around 700 - 50% of the total records) and the curve is more or less normally distributed.

```{r}
ggplot(data = df, mapping = aes(x = total_members, y = incoming_agricultural)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(title = 'Agricultural incomes vs expenditures')

# The number of people in the family doesn't relate that much to the agricultural income of their family
```
The Agricultural income of a family does not depend upon the number of people in that family.

```{r}
Depression <- as.character(df$depressed)
options(scipen=999)

ggplot(df, aes(x = factor(total_members), y = living_expenses, fill =Depression, color = Depression)) + 
  geom_bar(stat = "identity",position = "dodge")

```

As we have more records related to non-depressed people, we can see that living expenses are higher in their case (as its sums up more numbers). By intuiting, we can say that in most cases depressed people are spending lavishly (per person calculation) as they don't plan to spend on expenditures.

```{r}
ggplot(df,aes(x=Age,y=living_expenses,col=Depression))+geom_point()
options(scipen=999)
```


One thing we can analyze from the graph is that younger people have fewer living expenses and, as age increases, living expenses increase.

```{r}
boxplot(df$education_level ~ Depression, col = c("#FFE0B2", "#c57B00"))
```

In the above plot, we could see how education level values are distributed across two categories of our 'depressed' column. 

Interestingly, People who are depressed and those who are not depressed have the same mean(nearly). 

People in the non-depressed category have varying educational levels from 0 to 15 + that's why you can see some outliers on either side. Also almost most of them have education levels from 7-10

People in the depressed category moslty have education levels around 6-10 with mean close to 9 and only very few outliers on the right.

```{r}
hist(df$incoming_agricultural, prob = TRUE)                              
lines(density(df$lasting_investment), col = "#006400")
```

Agricultural incomes are a bit right skewed - some of the families are getting high profits and in general, most of the families are getting around 30 - 40,000k rupees.

```{r}
library(scales)
new <-df %>%
  group_by(Number_children) %>%
  summarise(n = n()) %>%
  mutate(Freq = n/sum(n))
new$Number_children <- as.character(new$Number_children)

geeks <- new$n
labels <- new$Number_children
 
piepercent<- round(100 * geeks / sum(geeks), 1)
 
# Plot the chart.
pie(geeks, labels = piepercent,
    main = "Children Percentage Pie Chart", col = rainbow(length(geeks)))
legend("topright", labels,
                    cex = 0.5, fill = rainbow(length(geeks)))

```

In the above plot, we could see how the number of childrens column is distributed - most families (around 71%) are having from 0-3 children with 1 and 2 children being occurred in most of the cases.



## Analysis 

Checking the correlations between the data attributes and with the target label

```{r}
library(corrplot)
df_pesronal <- df[c(3,4,5,6,7,8,23)]
df_financial <- df[c(9,12,14,21,23)]
corrplot(cor(df_pesronal), method = "number", type = "upper")
corrplot(cor(df_financial), method = "number", type = "upper")
```
Age has the highest correlation with the 'depressed' column, but as our target label is categorical, it is better to see the relationship in another way than just looking at the correlation values. 
Plot shows how different variables (Age and Gender) affect the target label.

```{r}
t <- df %>% select(depressed, Age) %>% group_by(Age) %>% summarise(count = n(), 
                                                                   avg_depression = mean(depressed))
#t$avg_age == 0.25
t[order(t$avg_depression, decreasing = TRUE),]
```
Depressed people are slightly of more age. Of all the age values which have an average depression greater than 0.25 (24 different age groups), the second least age group is 39 (i.e., 23 out of the top 24 cases have age greater than 39)

Interesting thing - the remaining 1 age group left in the top 24 groups is Age 17, which is the minimum age value in this data set.

```{r}
df %>% ggplot(aes(x = sex, y = Age, color = depressed)) +
  geom_point()
```

In males (category value 0), we can see that depressed people are distributed across the whole age group. In females, people belonging to the age group of 60 and more are more depressed comparatively.

```{r}
df %>% select(sex, depressed) %>% group_by(sex,depressed) %>% summarise(count = n())
```

Out of 115 males in the dataset, 20 of them are depressed (around 17%).
Out of the 1294 females in the dataset, 215 of them are depressed (around 16%)

Though there are more records of females, but from our analysis we could say that Gender doesn't have much effect on whether the person is depressed or not.

# Conclusion

After performing some analysis on whether and how the age and gender columns are going to affect depression, we can conclude that - 

1) People belonging to higher age groups tend to have more chances of being depressed(especially for females). However, a significant number of very young people have the potential to become more involved in these types of activities.We could observe a positive correlation overall.

2) Gender doesn't have a strong impact on depression.

## Any possible Sources of bias in the data set?

```{r}
df %>% count(sex)
df %>% count(Married)
df %>% count(incoming_salary)
df %>% count(depressed)
```

We can see that in all 4 columns listed above, one category completely dominates the other in terms of the number of times it has occurred.
Most of the data is that of females, married, and those who don't have any fixed income (salary), and mainly in the dependent column "depressed"—it has only 15% of records that give us information about depressed people.

So, we can say that this dataset is not a true sample, i.e., not a true representation of the whole population, and it does have a bias in it.


# GitHub Link

https://github.com/Charan-AB/Data-science-in-psychology